\section{Performance of the Compositional Algorithm}\label{sec:performance}

\newcommand{\exampleRef}[5][0]{%
\item[] {%
\begin{tabular}{r l}%
    \emph{system}:& {\bf #2} (\secref{sec:example-#3})\\
    \emph{initial}:& \parbox[t]{0.75\linewidth}{#4} \\
    \emph{target} $(\mathlarger{\mathlarger{\ifthenelse{#1=1}{\yesReachable}{\noReachable}}})$:& \parbox[t]{0.75\linewidth}{#5} \\
\end{tabular}}}

We now turn to examining the performance of \algref{alg:PNBAlgorithm}, using
the example systems introduced in \secref{sec:benchmarks}.  To do so, we must
first specify the particular initial and target markings that we will use for
each system. These markings are specified in \figref{fig:slowmarkings} (some
are expected to be reachable, some not).

\begin{figure}[ht]
\begin{multicols}{2}
\begin{itemize}
    \exampleRef[1]{\bufferSys{-}}{buffer}
        {``empty'': only tokens in upper places}
        {``full'': only tokens in lower places}
    \exampleRef{\tokenringSys{-}}{token-ring}
        {System ready: workers idle}
        {System started: \emph{all} workers working}
    \exampleRef[1]{\contree{-}{-}}{trees}
        {Root contains the only token}
        {Leaves each contain a token}
    \exampleRef{\distree{-}{-}}{trees}
        {Root contains the only token}
        {Leaves each contain a token}
    \exampleRef[1]{\cliqueSys{-}}{cliques}
        {Injector contains the only token}
        {Final (rightmost) place contains the only token}
    \exampleRef[1]{\powersetSys{-}}{powersets}
        {Injector contains the only token}
        {All places except root contain a token}
    \exampleRef{\overtakeSys{-}}{over}
        {All locks and interfaces unlocked, each car ready}
        {All cars overtaking}
    \exampleRef[1]{\hartstoneSys{-}}{hartstone}
        {Controllers, tasks and master all ready}
        {Each task working}
    \exampleRef[1]{\iteratedchoiceSys{-}}{iter-choice}
        {\addtokC{} contains the only token}
        {Alternating taken/not-taken marking, \addtokC{} has no token}
    \exampleRef[1]{\replicatorsSys{-}}{replicator}
        {\addtokC{} contains the only token}
        {\taketokC{} contains the only token}
    \exampleRef{\DACSys{-}}{dac}
        {Controller and workers ready}
        {Every worker and controller awaiting a join}
    \exampleRef[1]{\diningphilosophersSys{-}}{philos}
        {Philos ready, forks present}
        {Every fork taken}
    \exampleRef[1]{\cyclicschedulerSys{-}}{cyclic}
        {Injector contains token, each scheduler ready}
        {Each task working}
    \exampleRef[1]{\counterSys{-}}{counter}
        {Counter set to 0: all components zero}
        {Counter set to n: tester recognises counter is full}
\end{itemize}
\end{multicols}
\newcommand{\capt}{Example markings used in benchmarking
\algref{alg:PNBAlgorithm}.}
\caption[\capt]{\capt{} Those marked $\yesReachable$ are expected to be
    reachable and those $\noReachable$ are not}
\label{fig:slowmarkings}
\end{figure}

Given these marked examples, we execute our implementation, recording the total
time taken, averaged over 5 runs. The results are illustrated in
\tabref{tab:slowtimings}.

\input{results_timemem_slow}

Consider the first entry in \tabref{tab:slowtimings}, that of $\bufferSys{3}$,
which we evaluated by hand at the end of \secref{sec:compStatespaceGen}:
the maximum composition size encountered there (illustrated in
\figref{fig:exprtreestep3}) was $(8,1)$, agreeing with the recorded result.

For all systems, small parameters lead to run-times of $< 1$ second; but
small increments in the parameter sizes lead to vastly slower run-times, of
many seconds, accompanied with large composition sizes. Indeed, moving from a
parameter of $4$ to $5$ for $\overtakeSys{-}$ is catastrophic for the
algorithm's performance, with a much larger \TNFA{} composition encountered.

Indeed, the largest encountered compositions grows exponentially with the size
of the system - indeed, this should be intuitively clear, since the number of
markings of a net (and thus the number of states of its \TNFA{}) is exponential
in the number of places.

A final point to note is that in some cases, in particular,
\tokenringSys{-},\overtakeSys{-}, and \cyclicschedulerSys{-}, the result for
$\aN = 1$ differs to that for higher values of $\aN$. Indeed, inspecting the
desired markings for these systems, we observe that they specify the target
marking as a behaviour of all components, which can/cannot be reached for
multiple components, but however \emph{is} easily reachable for single
components.

Summarising these results, and to set the scene for the following chapter, we
can extract some key points that we will address:
\begin{enumerate}
    \item Without mitigation, the statespace explosion leads to NFA sizes that
        grow exponentially. We will investigate \emph{weak language preserving}
        reductions to reduce NFA sizes, while preserving correctness
        (\secref{sec:boundaryProtocol}).
    \item The behaviour of \bufferSys{3} should intuitively be no different to
        that of \bufferSys{4}, yet the current compositional algorithm does not
        exploit this. We will investigate \emph{fixed-points} of behaviour in
        such systems, and use \emph{memoisation} to avoid repeated work in
        their presence (\secref{sec:memoisation}).
    \item While PNB composition \emph{is} associative (up to isomorphism)
        w.r.t. the generated \TNFA{}, we will see that when employing
        reduction, it is \emph{not} associative w.r.t. the (size of)
        intermediate \TNFA{}s generated and thus the \emph{performance} of the
        algorithm. We will show that the performance is subtly affected by
        reassociating certain compositions (\secref{sec:reassociation}).
\end{enumerate}
